Hyper-linked Communications: WebRTC enabled asynchronous collaboration

### 2016-10-27 artigo/paper.tex
Initial version of paper to submitt to ICC 2017
Still needs to be shortened to 6 pages.
Working on the introduction.

\documentclass[conference,compsoc,a4paper]{IEEEtran}
\usepackage{geometry}
 \geometry{
 a4paper,
 total={170mm,257mm},
 left=20mm,
 top=20mm,
 }

\usepackage[backend=bibtex]{biblatex}
\usepackage{array}
\usepackage[utf8]{inputenc}   % <<<<< Linux
\usepackage[nomain,acronym,xindy,toc]{glossaries} 
\usepackage{graphicx}
\usepackage{tabularx} 
\usepackage{float}

\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}
%Includes "References" in the table of contents
%\usepackage[nottoc]{tocbibind}

\bibliography{references} % or

@@

\author{\IEEEauthorblockN{Henrique Rocha}
\IEEEauthorblockA{Instituto Superior T\'{e}cnico\\}
\IEEEauthorblockN{Av. Prof. Dr. Anibal Cavaco Silva\\
2744-016 Porto Salvo, Portugal\\
henrique.rocha@tecnico.ulisboa.pt}}

@@

\begin{abstract}
The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content. This paradigm allows to synchronize, structure and navigate communication content integrated into voice and video calls.

Voice and image together can express emotions like no other medium can. With hypermedia concepts, we can add more value to conference calls.

Our goal was the development of an application targeted to the web platform, resorting to \emph{WebRTC}, that leveraged the hyper-linked communications by providing a conference environment enriched with multiple media types, collaborative text editors, time annotations, instant messaging and a mechanism to superimpose hyper-content to video.
\end{abstract}

@@

\subsection{Background}
\label{section:background}

	As communications technologies appeared, we adapted the way we communicate. The purpose of this project is not the replacement of the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.

	With the advent of WebRTC and its successive integration with web browsers, it became possible to develop video conference web applications without plugins, this presents a range of possibilities on what can be implemented using already existing web technologies.

@@

\subsection{Outline}

This rest of this document is structured as follows:

@@

\item \textbf{Chapter \ref{chapter:relatedwork}} describes the previous work in the field.
\item \textbf{Chapter \ref{chapter:architecture}} describes the system requirements and the architecture for an Web Application that fulfills the goals of this thesis.
\item \textbf{Chapter \ref{chapter:implementation}} describes the implementation of our Web Application and the technologies chosen.
\item \textbf{Chapter \ref{chapter:evaluation}} presents the evaluation tests performed and the corresponding results.
\item \textbf{Chapter \ref{chapter:conclusion}} summarizes the work developed and proposes future work.
\end{itemize}

One way to overcome the \gls{IPv4} address scarcity problem \cite{ipv4} was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \gls{IP} address with a different port. This technique is known as \gls{NAT}\cite{rfc3489}.

\gls{NAT}s weaknesses are being exposing at the application layer, namely impacting applications that require direct communications between two private networks.

In order to implement an hyper-linked communication solution, several design decisions had to be made. The limitations imposed by the \emph{Internet's} structure, its protocols and a browser's capabilities are key factors to consider when implementing a solution that allows bi-directional communications, interactive media and collaboration environment.

Due to the use of \gls{NAT}, bi-directional communications between clients have different needs from request-response based communications between clients and servers.
This lead to the appearance of mechanisms such as \gls{STUN}, \gls{TURN} and \gls{ICE}, in order to bypass the limits imposed by \gls{NAT}.

\gls{STUN}, \gls{TURN} and \gls{ICE} \cite{natvoip} servers are a possible solution to overcome \gls{NAT}'s exposed weaknesses to applications that require communications between two private networks.

When connection is established, \gls{WebRTC} came to simplify how audio and video are transmitted through web browsers. \gls{WebRTC} is an open source technology that defines a collection of standard protocols and \emph{JavaScript} \gls{API}s for web browser based real time communications without installing any additional application or plug-in.

@@

  One of \gls{WebRTC} signaling's requisites is bi-directional communication. The \emph{WebSocket} protocol allows bi-directional communications over a full-duplex socket channel \cite{rfc6455}, by other words it supports sending and receiving data simultaneously.

With the communications establishment issue solved, we had to discuss the different types of media and what can be done with each kind in order to increase the value of communications among users. In this context, we have studied solutions and libraries that allow us to implement our prototype with time manipulation features, collaborative text edition, record and playback interactive video.

@@

  Using technologies that relies only on web standards, like \gls{CSS}, \gls{HTML}5, \emph{JavaScript} and \gls{SVG}, will make possible to develop an application that applies the hypermedia concept with the advantage of being compatible with a greater amount of web browsers.

@@

@@

\subsection{Requirements}
In a general way our system's goal is to provide a multi-party video and audio conference environment that supports chat, time manipulation, collaborative text edition and hyper-content creation.

@@

%Sets the bibliography style to UNSRT and imports the 
%bibliography file "samples.bib".
%\bibliographystyle{unsrt}
%\bibliography{references}

%\input{thesis.bbl}
\printbibliography
\end{document}

\grid
\grid

>>>>

%\documentclass[conference,compsoc,a4paper]{IEEEtran}
\documentclass[10pt,conference]{IEEEtran}

\usepackage{array}
\usepackage[utf8]{inputenc}   % <<<<< Linux
\usepackage[nomain,acronym,xindy,toc]{glossaries} 
\usepackage{tabularx} 
\usepackage{float}

\usepackage{cite}    
\usepackage{graphicx,flafter}
\graphicspath{{../tese2/}}

\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

\usepackage{url}

@@

\author{\IEEEauthorblockN{Henrique Rocha}
\IEEEauthorblockA{INESC-ID / Instituto Superior T\'{e}cnico\\
Av. Prof. Dr. Anibal Cavaco Silva\\
2744-016 Porto Salvo, Portugal\\
Email: henrique.rocha@tecnico.ulisboa.pt}
\and
\IEEEauthorblockN{Ricardo Lopes Pereira}
\IEEEauthorblockA{INESC-ID / Instituto Superior T\'ecnico \\
Av. Prof. Dr. Cavaco Silva\\
2744-016 Porto Salvo, Portugal\\
Email: ricardo.pereira@inesc-id.pt}
}

@@

\begin{abstract}
The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, to the communication realm.
This paradigm allows to synchronize, structure and navigate communication content integrated into voice and video calls.
Voice and image together can express emotions like no other medium can. With hypermedia concepts, we can add more value to conference calls.

This paper describes the design, implementation and evaluation of a hyper-linked communication platform targeting the web platform and making use of \emph{WebRTC} technology.
Our prototype provides a hyper-linked, non-linear, multi-party video conferencing platform with support for multiple media types, collaborative text editor, time annotations, instant messaging and a mechanism to superimpose hyper-content to video.

Our prototype was built using a hybrid peer-to-peer and client-server architecture, resorting to open source software.
Usability tests using 20 users evaluated our prototype positively: 100\% considered it innovative and 95\% recommended its use. 
From a deployment point of view, our performance tests showed that our solution was expensive to scale up, even though the cost could be significantly lower by disabling some of the functionality.

\end{abstract}

@@

As communications technologies appeared, we adapted the way we communicate. 
Multi-party video conference is today a commonly used tool in the enterprise, academia and personal live.
The \gls{WWW} has made us familiar with the concept of text hyper-media, in the form of hyper-links.
\gls{IPTV} and video hosting platforms, such as Youtube, made non-linear video watching a common practice.
The later also use the concept of hyper-links superimposed over videos.
However, these three concepts, video-conferencing (which is real-time), non-linear video watching and hyper-media, have largely remained independent concepts.

The purpose of this project is not the replacement of the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.

The advent of \gls{WebRTC}, currently available on the major web browsers, made it possible to develop video conference web applications without plugins, greatly increasing the range of what can be implemented using web technologies.

@@

The remainder of this paper is organized as follows.
Section \ref{chapter:relatedwork} provides an overview of previous work in the field.
Section \ref{chapter:architecture} describes the requirements we defined for our system and architecture for fulfilling them.
The choices performed for the implementation of our prototype are presented in Section \ref{chapter:implementation}.
In Section \ref{chapter:evaluation} we present the evaluation tests performed and their results.
Section \ref{chapter:conclusion} summarizes the work developed and proposes future work.

Since the early days of video technology, one of the problems raised consisted on how to add more information onto video without generating multiple versions. This section examines technologies that allows different ways to present multimedia content in such a way that it change based on synchronization amongst other multimedia elements or user interaction.

\emph{Hypertext} is a type of text that provides links to texts or other types of content, these links are known by \emph{hyperlinks}. \emph{Hypermedia} is an evolution of \emph{hypertext}, it includes audio, images, text and video.

\emph{Hypermedia} concept brings the possibility to organize and overlay multimedia elements into a nonlinear linear structure.

In the beginning of analog video technology, the navigation over it was quite limited to simple operations such as play, stop, rewind and fast-forward. As video started to be digitalized, new operations over video emerged, such as random jumps and chapter navigation through interactive menus present on \gls{DVD}.

%RP depois disto falta uma introdução a explicar que precisas de controlar a apresentação e execução de diversos conteúdos e que para isso vais precisar de uma tecnologia que permite sincronizar e controlar. Esta secção examina isso.

Some \gls{MPEG} implementations, like \cite{embedded}, added hypermedia information to empty space present on \gls{MPEG} frames in order to provide interactive television, modifying the \gls{MPEG} encoder and decoder in order to handle hypermedia content.
%RP esta última frase, com detalhes de implementação fica algo deslocada antes de apresentares os exemplos de hypermedia. Talvez deslocar para mais tarde.
Hypermedia is a concept that holds the promise of future technology and features but it is also already present in our daily lives.
%RP por exemplo na publicidade no youtube

  Subtitles are an example of information that might be required.
  The need to translate movies, raised the problem whether it is appropriate to change the original video or audio. For example, subtitles should be an entity independent from the video, in order to be personalized or replaced easily.

  \gls{SAMI}, and \gls{SRT} are two of the multiple formats for subtitles commonly supported by video players. Although those formats have styling available, they are quite limited to text.

  Hyper-video is a kind of video that contains links to any kind of hypermedia, including links to skip part of it. An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time in a video or audio track. \emph{HyperCafe}~\cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted of an interactive film that enabled switching between different conversations taking place inside a cafe.

  Detail-on-demand is a subset of hyper-video that allow us to obtain additional information about something that appears along the video, like obtaining information about a painting that appears in a particular segment. \emph{Hyper-Hitchcock}~\cite{hitchcock} is an editor and player of detail-on-demand video.

  In order to navigate through a dynamic video, we must be aware of time synchronization and the multiple time flows, it ils important that all time, causality and behavior rules are well defined.
  %RP one must be aware? O utilizador? Ou queres dizer que é preciso definir as relações? O que falas a seguir é sobre linguagens para isso.

  \emph{HyVAL}~\cite{hyval} is an \gls{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia. HyVAL defines defines video structure, internal video and external media objects. 
  \emph{HyVAL} uses a primary video stream, around which all other elements are organized and synchronized.
  \emph{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically. 
  This approach is quite restrictive if we want to apply hyper-video concepts to videos that do not follow this structure. External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.
  %RP não percebo bem a última frase. O que é primary video?

  \gls{SMIL}~\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films. With \gls{SMIL} it is possible to synchronize multiple videos, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.
  %RP synchronize multiple sections of video? vários videos apresentados em simultâneo?

  \gls{SMIL} is an \gls{XML} based language that defines twelve modules: \emph{Animation}, \emph{Content Control}, \emph{Layout}, \emph{Linking}, \emph{Media Objects}, \emph{SmilText}, \emph{Meta Information}, \emph{Structure}, \emph{Timing}, \emph{Time Manipulations}, \emph{State} and \emph{Transitions}.

@@

  \item The \textbf{Animation} module contains elements and attributes that define a time based mechanism for composing the effects of animations. For example, this module can perform changes on \gls{XML} or \gls{CSS} attributes like color and dimensions.

  \item The \textbf{Content Control} module contains elements and attributes that provide optimized alternatives for content delivery. For example, it could be used to change audio language in function of user's nationality, for videos with multiple audio channels.

  \item The \textbf{Layout} module contains elements and attributes for coloring and positioning media content. Other layout mechanisms are also possible, such as \gls{CSS}.

  \item The \textbf{Linking} module contains elements and attributes for navigational \emph{hyperlinking}. Navigation can be triggered by events or user interaction.

  \item The \textbf{Media Object} module contains elements and attributes for referencing rendering behavior of external multimedia or control objects.

  \item The \textbf{SmilText} module contains elements and attributes that define and control timed text. For example, this module could be used to create labels and captions.

  \item The \textbf{Meta Information} module contains elements and attributes that allows describing the \gls{SMIL} document. For example, this module could be used to define movie details such as category, director, writers and cast.

  \item The \textbf{Structure} module defines the basic elements and attributes for structuring \gls{SMIL} content. This module defines a \emph{head} element that contains non temporal behavior information defined by  \emph{Meta Information}, \emph{Layout} and \emph{Content Control} modules. This module also defines the \emph{body} element, where all temporal related module information is contained.

  \item The \textbf{Timing} module is the most important module on \gls{SMIL} specification. Due to its complexity, it is divided into seventeen sub-modules for coordination and synchronization of media over time. The three main elements are \emph{seq}, \emph{excl} and \emph{par}, that, respectively, play child elements in sequence, one at a time and all at the same time.

  \item The \textbf{Time Manipulations} module adds time behavior attributes to \gls{SMIL} elements, such as speed, rate or time.

  \item The \textbf{State} module defines attributes that define the state of \gls{SMIL} elements, such as element visibility, current element time, amount of repeated loops, playing state and many others.

  \item The \textbf{Transitions} module defines attributes and elements for transitions across multiple \gls{SMIL} elements according to the \emph{Timing} module.

The \gls{DOM} is a standard \gls{API} that allows easy management of documents that are organized in a tree structure, by providing \gls{CRUD} operations over its elements and their attributes. \gls{DOM} makes it easy to inter-operate between imperative and declarative programming languages~\cite{dom}.
%RP não arranjas uma referência para o dom?

Like \gls{DOM}, \gls{SMIL} \gls{DOM} is an \gls{API} for \gls{SMIL} documents. Allowing \gls{CRUD} operations over \gls{SMIL} documents is an important feature for extending \gls{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \emph{JavaScript} functions.

  \gls{SMIL}'s modules can be used to synchronize and animate \gls{XHTML} and \gls{SVG} elements.
  %RP ``are used'' ou ``can be used''?

  \gls{SMIL} fits our goals for creating a multimedia rich hyper-call, but it lacks on browser compatibility. Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \gls{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plug-in.

  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \gls{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, does not require a plug-in to be installed and should not have incompatibility issues. 
  SmillingWeb already takes advantage of \gls{HTML}5 and \gls{CSS}3.
  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}(accessed June 2, 2015).}, a simple \emph{JavaScript} library that may require plug-ins if new features are not supported.  
  But SmillingWeb just implements a subset of \gls{SMIL} 3.0 and their scheduler engine loads the \gls{SMIL} file only once, which could raise problems when dealing with \gls{SMIL} changes due to real time communications.
  Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to high latency networks leading to  pauses during playback.

% XXX
  %RP fazes uma mudanças de tema muito bruscas. Podes usar \subsubsection ou \paragraph para dividir isto melhor. E uma frase entre temas a servir de ``cola'' ou ``separação'' também não fica mal

  An alternative to \gls{SMIL}'s \emph{Layout} module is to use \gls{HTML} which is a markup language based on \gls{XML} that is used for creating web pages. \gls{HTML} alone is a very poor language when we are focused on visual appealing and interactive web pages. Languages like \gls{CSS} and \emph{JavaScript} are typically used along with \gls{HTML} for improving the interaction and appearance of a web page.

  %RP esta parte está um pouco confusa. Primeiro falas de smil, depois de html, depois de smil novamente. Devias agrupar tudo. Devias agrupar as soluçoes por tecnologia. Se preciso, primeiro fala de html, css, etc, e depois de smil e das implementação de smil (incluindo as que são implementadas em html e js).

  \gls{CSS}'s goal is to separate the structure of an \gls{XML} document from its appearance. \gls{CSS} defines styles for \gls{XML} tags based on their name, class, identifier or position.
  Besides static styling \gls{CSS} also supports animation and transitions leading to more dynamic content.
  %RP Tens muitas alturas em que não começas frases novas. Usas virgulas para separar conteúdo que não faz sentido ficar na mesma frase pois começa com um novo sujeito seguido de verbo.

  \emph{JavaScript} is an imperative object-oriented language based on \emph{ECMAScript}. It is used mainly on client-side and executed by a web browser. \emph{JavaScript} has its own implementation of \gls{DOM} and one of its advantages is the ability to download and execute code on-the-fly without the need of pre-installed plug-ins.

  \emph{JavaScript} has compatibility issues among the different web browsers, leading to different behaviors. To solve that problem, there are libraries written in \emph{JavasSript}, namely \emph{jQuery}, that implements the same functionality for multiple browsers, masking most of the incompatibility issues.

  With the emergence of \gls{HTML}5, tags like \emph{video}, \emph{audio} and \emph{track} allow us to play video with multiple \emph{codecs}, audio and subtitles in \gls{WebVTT} format. Another important tag is \emph{canvas} that allows drawing graphics with \emph{JavaScript} on a rectangle within a web page.

  For example, with \gls{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}(accessed June 2, 2015).}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call. Another example would be a collaborative spreadsheet using \gls{WebRTC}. With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.

  \gls{SVG} is an \gls{XML} based format that incorporates the animation module of \gls{SMIL}. Currently, \gls{SVG} allows adding movement and animating attributes of elements. When embedded on \gls{HTML}5, it allows dynamic changes to inner content in real time through the \gls{DOM} \gls{API}. Besides that, it also allows calling \emph{JavaScript} functions on events such as animation end, mouse over and mouse click.

  Back in 1995, \emph{flash}\footnote{\url{http://www.adobe.com/products/flash.html}(accessed June 2, 2015).} was developed for web-based animations. Introducing video support in 2002, \emph{flash} started to grow after that. Competitor's players, at that time were focused on playing video and audio, while \emph{flash} had vector graphics and focused on streaming \emph{on-demand} video across multiple platforms. \emph{VP6} was their choice of video \emph{codec}, providing half the video size (in Byte) for the same quality and providing video quality adjusted to \emph{Internet} connection latency. 
  In 2010, \emph{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \gls{HTTP} and streaming using \gls{RTMP}.

  \gls{RTMP} is a \gls{TCP} based protocol used for streaming audio, video and data between a \gls{FMS} and \emph{flash} players. A bidirectional connection is established between the two in order to allow real time communications. A \emph{flash} player can stream a webcam video to a \gls{FMS} using \gls{RTMP} or it can request a video stream to \gls{FMS} that can either be a pre-recorded stream, live stream or data. Multiple \gls{FMS} servers can be used in parallel to increase capacity and handle more streams simultaneously.
  %RP chained ou usados em paralelo?

  \gls{FMS} can stream video and audio to one or more subscribers by sending a separate copy for each subscriber. With \gls{RTMFP} it is possible to stream video directly between \emph{flash} players, allowing a publisher to break up a stream into pieces that can be cooperatively distributed in a P2P mesh. \gls{RTMFP} uses \gls{UDP} to speed packet delivery, which although it is not reliable, is well suited for video streaming. Like \gls{WebRTC}, \emph{flash} players also need to apply techniques like \gls{STUN} and \gls{TURN} for \gls{NAT} traversal.

  Although \gls{HTML}5, \emph{JavaScript}, \gls{CSS} and \gls{WebRTC} implement some of \emph{flash}'s features, it does not mean that \emph{flash} will be replaced soon.
  Instead, both technologies can be used to develop rich \emph{Internet} applications.
  It is also important to note that \gls{HTML}5 is better supported in mobile devices than \emph{Adobe Flash}.

  Like \emph{Flash}, \emph{Microsoft Silverlight}\footnote{\url{http://www.microsoft.com/silverlight/}(accessed June 2, 2015).} is a cross browser plug-in and platform that is used to develop rich \emph{Internet} applications. It supports vector graphics, animation and video. Compared to \emph{flash}, which uses \emph{ActionScript}, \emph{Silverlight} applications can use languages like C\#, \emph{VisualBasic} and \gls{XAML}. \emph{Silverlight} uses a technique called \emph{Smooth Streaming} from \emph{IIS Media Service} that consists on delivering video in real time with adjusted quality in function of bandwidth variations and \gls{CPU} usage.

  Using technologies that relies only on web standards, like \gls{CSS}, \gls{HTML}5, \emph{JavaScript} and \gls{SVG}, will make possible to develop an application that solves our problem with the advantage of being compatible with a greater amount of web browsers.

\gls{WebRTC} is an open source technology that defines a collection of standard protocols and \emph{JavaScript} \gls{API}s for web browser based real time communications without installing any additional application or plug-in.

@@

@@

@@

\subsection{Streaming and Recording}\label{recstream}~\\
	If, for instance, one wants to rewind a real time video, recordings will be needed from whom is streaming the video. 
 Our first concern on real time collaboration applications, besides the communication itself, is the data storage and representation. Storing multimedia content is not a viable solution because most browsers recommend limiting local storage to at most five megabytes per origin\footnote{\url{http://www.html5rocks.com/en/tutorials/offline/quota-research/} (accessed May 13, 2016)} .

 	In order to provide a way to record and playback streams, additional servers will be required to process and record the large amount of data generated by audio and video streaming.

 \gls{RTP}\cite{rfc3550} is used for streaming audio and video over \gls{IP}.
 Multimedia content is transported on the payload of \gls{RTP} messages, that contain headers for payload identification. \gls{RTP} is independent from its payload type, allowing it to transport any kind of encoded multimedia. A sequence number is used for sorting received packets.

\gls{RTCP} is used for controlling \gls{RTP} multimedia streams, it provides bandwidth statistics and control information that can be used for changing the quality of the stream in real time.

  \gls{RTP} allows to change its requirements and add extensions to it with profiles. One of the most used ones is the \gls{RTP} profile for audio and video \cite{rfc3551}, which lists the payload encoding and compression algorithms. This profile also assigns a name to each encoding which may be used with other protocols like \gls{SDP}.
  Another profile for \gls{RTP} is defined by \gls{SRTP}, which provides encryption, authentication and replay protection for \gls {RTP} traffic. The analogous secure protocol for \gls{RTCP} is \gls{SRTCP}.

  Both \gls{SRTP} and \gls{SRTCP} use \gls{AES} by default, which is a symmetric-key algorithm for data encryption. Each packet is encrypted using a distinct key-stream, as otherwise, using a single key-stream with \gls{AES} on \gls{CBC} would make it impossible to recover from packet loss.

  Two key-stream generators for \gls{AES} were defined: \emph{Segmented Integer Counter Mode} and \emph{f8-mode}. If a packet is lost, there is no impact on other packets, as the initialization vector is obtained through those key-stream generators and it is fixed for each packet.
  %RP isto está tudo no RFC3550? Se não, faltam muitas citações.
  %HR esse RFC é grande

  \gls{RTP} recorders are independent of payload encoding, they do not decode \gls{RTP} packets, they record packets instead, allowing to record all video and audio formats even if they're encrypted.

        %RP lá vais tu mudar de tema da conversa sem explicar porquê....
        Even though one on one calls are common, there are occasions when several people take part of the same video call.
	Multi-party video calls can be achieved on \gls{WebRTC} by streaming video from each participant to all the other participants. Although this works, the bigger a conference room is, the bigger is the used bandwidth to stream video to all participants within the conference room.
        In this scenario, a more efficient alternative to peer-to-peer is the use of a \gls{MCU}.

	\emph{Jitsi Video Bridge} \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}(accessed June 2, 2015).} receives one stream from every participant on a conference, either from a \emph{jitsi} client or a \emph{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends. Although all the participants still need to download all the streams from the \emph{Jitsi Video Bridge} server, typically download rates are much bigger than upload rates, making this solution more feasible.
	\emph{Jitsi Video Bridge} uses \gls{XMPP} as a signaling protocol and its \emph{colibri} extension \cite{xep0340} to reserve channels for video transmission. Despite this choice for signaling protocol, \emph{Jitsi Video Bridge} also supports \gls{SIP}.

	  \gls{KMS} supports trans-coding, group communications, recording, mixing, broadcasting, applying filters and performing image and sound analysis (for example it allows face and QRcode detection). \gls{KMS} functionalities are exposed through \emph{JSON-RPC} over \emph{WebSocket}.
          There are three clients available: \emph{JavaScript} client for web browsers, \emph{Java} client for \emph{Java EE} servers and \emph{JavaScript} client for \emph{Node.js} servers. 
	\gls{KMS} supports streaming over \emph{WebRTC}, \gls{HTTP} and \gls{RTP} endpoints. Another important component of \gls{KMS} is \emph{Kurento Repository}, which supports recording and playing directly from \emph{MongoDB}. That is important for providing a scalable media storage. Unlike \emph{Jitsi Video Bridge}, \gls{KMS} does not enforce a specific signaling protocol.

@@

\subsection{Requirements}
\label{chapter:requirements}
In a general way our system's goal is to provide a multi-party video and audio conference environment that supports chat, time manipulation, collaborative text edition and hyper-content creation.

@@

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

\bibliographystyle{IEEEtran}
\bibliography{../tese2/references}

\end{document}
